(An introduction to )MultiSymplectic Geometry and Classical Field systems

### 2019-03-01 cps-via-ms.tex
small fix

\item It turns out that  $\LinDualJet $ carries a naturally defined multi-symplectic form $\omega = -\ExtD \theta$ derived from a multicanonical $m$-form $\theta$, obtained by pulling back the tautological $m$-form $\theta_\Lambda$ from $\bigwedge^m (E)$.\\
	Thus, it can be considered the theoretical analogue of the cotangent bundle in ordinary geometric mechanics.

>>>>

\item It turns out that  $\AffDualJet $ carries a naturally defined multi-symplectic form $\omega = -\ExtD \theta$ derived from a multicanonical $m$-form $\theta$, obtained by pulling back the tautological $m$-form $\theta_\Lambda$ from $\bigwedge^m (E)$.\\
	Thus, it can be considered the theoretical analogue of the cotangent bundle in ordinary geometric mechanics.

### 2019-03-01 ms-mechanics.tex
small fix

\title{(An introduction to )MultiSymplectic Geometry and Classical Field systems}
\author{
	International Ph.D. Program in "Science" \\
	"Differential geometry and applications to modern physics" \\
	Department of Mathematics and Physics\\
	Universit\'a Cattolica del Sacro Cuore\\
	via Musei 41, 25121 Brescia, \underline{Italy}
}

\documentclass[a4paper,12pt,fleqn]{scrartcl}  %Per La Stampa
\usepackage[a4paper, margin=2cm]{geometry}

@@

\begin{abstract}
$n$-plectic structures (also called \emph{multisymplectic}) are a rather straightforward generalization of symplectic ones where closed non-degenerate $n+1$-forms take the place of  $2$-forms.
\\
As the same way as one can relate a symplectic manifold to any ordinary classical mechanical system (e.g. a single point-like particle constrained to some manifold), it is possible to associate a multisymplectic manifold to any classical field systems (think them as a continuum medium as a filament or a membrane).
\\
The aim of my talk is to give an account on the multisymplectic framework for (I-order) classical fields theories comparing it with another object that plays a significant role in the mathematical description of classical fields called \emph{Covariant phase space}.
\\
Being the latter a sort of "$\infty$-dimensional manifold" (namely a mapping space), we will from this picture the idea that multisymplectic geometry could be seen as a tool that allow us to treat such formal object in a finite dimensional setting.

@@

\subsection{Far capire cosa intendo per campo classico}
Classicamente: DOF come numero di misure necessarie per fissare univocamente una configurazione.
\begin{table}[]
\begin{tabular}{lll}
	System & Configuration space (locally) & protoype 
 	\\
	Ordinary system & $\mathbb{R}^n$ with $n\in \mathbb{N} = Hom_{set}(\{1,\ldots n\}, \mathbb{R})$ & A point on a $n$-dimensional mfd.
	\\
	1 dimensional field & $\mathbb{R}^\mathbb{R} = C^\infty(\mathbb{R},\mathbb{R})$ & collection of points on a line each one parametrized by ...
	\\
	n dimensional field & $(\mathbb{R}^m)^{\mathbb{R}^n} = C^\infty (\mathbb{R}^n, \mathbb{R}^m)$ & ...
\end{tabular}
\end{table}

\subsection{spunti Open problems}
	We have a momap on ms mfd and a momap on mapping space (as diffeological space) does the two match?

	We have a canonical m-s structure on J.. and a canonical poisson structure con cov phase space (Peierls bracket) does the two match?

\section{Spunti dei Saggi}
\subsection{Rogers: arXiv:1009.2975v3}
Multisymplectic manifolds naturally arise in certain covariant Hamiltonian formalisms for classical field theory.
In these formalisms, one describes a $(n+1)$-dimensional field theory by using a finite-dimensional $n$-plectic manifold as a \emph{multi-phase space} instead of an infinite-dimensional phase space. 
The $n$-plectic form can be used to define a system of partial differential equations which are the analogue of Hamilton's equations in classical mechanics. 
The solutions to these
equations correspond to particular submanifolds of the multi-phase space that encode the value of the field at each point in space-time as well as the values of its time and spatial derivatives.

>>>>

\title{(An introduction to )MultiSymplectic Geometry and Classical Field systems}
\author{
	International Ph.D. Program in "Science" \\
	"Differential geometry and applications to modern physics" \\
	%
	\begin{minipage}[c]{0.5\textwidth}
		\vspace{2em}
		\centering
		Department of Mathematics and Physics\\
		Universit\'a Cattolica del Sacro Cuore\\
		via Musei 41, 25121 Brescia, \underline{Italy}
	\end{minipage}
	\hfill
	\begin{minipage}[c]{0.5\textwidth}
		\vspace{2em}
		\centering
		Department of Mathematics\\
		KU Leuven\\
		Celestijnenlaan 200B, 3001 Leuven, \underline{Belgium}
	\end{minipage}
}

\documentclass[a4paper,12pt,fleqn]{article}  %Per La Stampa
\usepackage[a4paper, margin=3cm]{geometry}

@@

$$
\ast 
\star 
\oast 
\ostar
\smoast
\smostar
J^\smostar
J^\smoast
J^\ostar
$$

\begin{abstract}
$n$-plectic structures (also called \emph{multisymplectic}) are a rather straightforward generalization of symplectic ones where closed non-degenerate $n+1$-forms take the place of  $2$-forms.
\\
Just as one can associate a symplectic manifold to an ordinary classical mechanical system (e.g. a single point-like particle constrained to some manifold), it is possible to associate a multisymplectic manifold to any classical field systems (e.g. a continuous medium like a filament or a membrane).
\\
The aim of this talk is to give an account on the multisymplectic framework for (I-order) classical fields theories trying to comparing it with another object that plays a significant role in the mathematical description of classical fields called \emph{Covariant phase space}.
\\
Being the latter a sort of "$\infty$-dimensional manifold" (namely a mapping space), we will draw from this picture the idea that multisymplectic geometry could be seen as a tool that allows us to treat such formal object in a finite-dimensional setting.

@@

\section{A glance at the geometric approach to classical mechanics}
Recall: Geometric Mechanics of Classical system
Configuration space
Generalized velocities space and phase space
...

\section{Far capire cosa intendo per campo classico}
Classicamente: DOF come numero di misure necessarie per fissare univocamente una configurazione.
\begin{table}[]
\begin{tabular}{lll}
	System & Configuration space (locally) & protoype 
 	\\
	Ordinary system & $\mathbb{R}^n$ with $n\in \mathbb{N} = Hom_{set}(\{1,\ldots n\}, \mathbb{R})$ & A point on a $n$-dimensional mfd.
	\\
	1 dimensional field & $\mathbb{R}^\mathbb{R} = C^\infty(\mathbb{R},\mathbb{R})$ & collection of points on a line each one parametrized by ...
	\\
	n dimensional field & $(\mathbb{R}^m)^{\mathbb{R}^n} = C^\infty (\mathbb{R}^n, \mathbb{R}^m)$ & ...
\end{tabular}
\end{table}

A questo punto si dovrebbe Essere chiaro il punto di partenza:
\\

\section{Kinematics of Classical fields}
The starting point is the \emph{configuration bundle} 
 	\begin{displaymath}
 		E = \left( \pi_{M,E} : E\rightarrow M \right)
 	\end{displaymath}
is a smooth bundle encoding the kinematics.

\subsection{Jet Bundle}
The first jet bundle of $E$ generalises the space of generalized velocities.

The fiber over a point $y\in E$ is
\begin{displaymath}
	\left( J^1 E\right)_y = \left\lbrace \gamma \in \text{L}(T_{\pi(y)}M, T_y E) \; \vert \: T\pi \circ \gamma = \text{id}_{T_{\pi(x)} M} \right\rbrace
\end{displaymath}
(T = Tangent functor L = space of linear maps)

which is and affine subspace of $\text{L}(T_{\pi(y)}M, T_y E)$ modelled on the following (difference) vector space:
\begin{displaymath}
	\vec{\left( J^1 E\right)_y} = \left\lbrace z \in \text{L}(T_{\pi(y)}M, T_y E) \; \vert \: T\pi \circ z = 0 \right\rbrace
	\cong text{L}(T_{\pi(y)}M, T_y E)
\end{displaymath}

\section{Spunti / Open problems}
\begin{itemize}
	\item We have a momap on ms mfd and a momap on mapping space (as diffeological space) does the two match?
	\item We have a canonical m-s structure on J.. and a canonical poisson structure con cov phase space (Peierls bracket) does the two match?
	\item there are other notion of moment map on the multisymplectic setting. For instance how the definition in section 4 of \cite{Gotay1998a} is related to the one in \cite{Ryvkin2018}?
\end{itemize}

\nocite{*}
\bibliographystyle{alpha}
\bibliography{biblio}
